### 2013-11-12 paper.tex
committing an intermediate state on the ride home...

\documentclass[preprint,9pt]{sigplanconf}
\usepackage{url}
\usepackage{listings}
\usepackage{subfigure}
\usepackage{times}
\usepackage{multirow}
\usepackage{color}
\usepackage{multicol}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{color}
\usepackage{rotating}

\newcommand{\addtodo}[1]{\textcolor{red}{[To do: #1]}}
\newcommand{\ctrap}{CTrap\xspace}
\newcommand{\ctraps}{CTraps\xspace}
\newcommand{\ctrapsfull}{CTraps-Full\xspace}
\newcommand{\ctrapsmm}{CTraps-NRR\xspace}
\newcommand{\lws}{CTraps-LWS\xspace}
\newcommand{\lwt}{LWT\xspace}

\newcommand{\Caption}[1]{\begin{minipage}{1.0\columnwidth} \caption{#1} \end{minipage}\vspace{-2ex}}
\newcommand{\CaptionWide}[1]{\caption{#1}\vspace{-2ex}}
%\doublespacing
\authorinfo{Anonymous for Submission}{}

\title{Data Provenance for Concurrent Program Debugging and Analysis in Production Systems}

%\authorinfo{Brandon Lucia \and Luis Ceze}{University of Washington, Department of Computer Science and Engineering}{\{blucia0a,luisceze\}@cs.washington.edu\\http://sampa.cs.washington.edu\\{\bf In Submission --- Please Do Not Distribute}}

%\author{Anonymous for Submission}
\begin{document}
\maketitle

\begin{abstract}
%Monitoring inter-thread memory dependences in production is useful.  Doing so
%improves debugging information and provides information to dynamic analyses and
%runtime systems.  Prior work that continuously monitors cross-thread
%dependences typically has overheads too high for production.  
We propose {\em last writer slices}, an analysis for shared-memory
multi-threaded programs that tracks {\em data provenance information} in
production code.  Last writer slices dynamically record the thread and
operation that last wrote each variable, reflecting the {\em provenance} of
values.  Provenance helps understand bugs and we provide last writer slices to
programmers during debugging.  We build {\em communication traps}
(\ctraps) using last writer slices.  \ctraps efficiently
monitors and can interpose on operations that share data between
threads.  \ctraps is extensible, exposing inter-thread dependences
to {\em \ctraps applications} that implement arbitrary analyses.

We implement last writer slicing and \ctraps.  We conduct debugging studies of
real, buggy programs showing last writer slices are useful for debugging.  We
show \ctraps is useful by implementing two concurrency analyses from prior
work.  We evaluate the performance of our designs using server programs
and standard benchmarks.  We show that collecting last writer slices
has low run time overheads (0--15\%) for many applications,including  
memcached, LevelDB, MySQL, and Apache.  We show that \ctraps imposes low 
overheads by default and that overheads scale with analysis complexity.

\end{abstract}

\section{Introduction}
Multi-threaded programming is challenging.  In contrast to the simplicity of
sequential reasoning, multi-threaded programs require complex reasoning about
many threads of control and their interactions.  In shared memory programs,
threads interact by reading and writing memory locations.  Threads' reads and
writes interleave arbitrarily and nondeterministically and each interleaving
can yield different behavior.  Reasoning about thread interactions is a central
challenge to building multi-threaded software because some thread interactions
have undesirable consequences, like causing a crash or degrading performance.

%A common approach to this challenge is to reason about dependent memory
%operations.  Memory operations are dependent if their order determines the
%values they read or write.  The utility of dependence information is made
%especially clear by its prevalent use during manual program debugging and in
%support of program analysis tools.

Debugging a concurrent program requires understanding how threads interact and
{\em data provenance information} furnishes that understanding. A value's
provenance is a description of the reason that it is stored in a variable.
Provenance can take on many forms, for example, a value's provenance could be
the instruction that wrote it or the function that returned it.  Provenance
information helps answer the question: "Why does this variable contain this
value?", which is central to debugging.

%During debugging provenance helps a programmer better understand not just where failures occur, but where values came from, connectng disparate
%parts of a program that interact, leading to a failure.

Collecting debugging information, like data provenance, from production systems
is particularly important for concurrent software.  Concurrency bugs manifest
as failures rarely.  When a production system fails and a programmer receives a
bug report, reproducing the reported failure to collect provenance information
may be difficult or impossible.  It is essential to collect as much information
as performance margins permit from production, to reduce the need for
time-consuming failure reproduction.

Prior work has previously shown that some limited forms of provenence help with
debugging~\cite{badapples, whylineicse, tipslicingsurvey}.  One effort tracked
code that wrote unusable values into variables~\cite{badapples}.  That work
showed that revealing the connection between an unusable value's origin and a
failure it causes is helpful during debugging.  Other work showed that
programmers debug more effectively when they can ask "why" questions about
values during debugging~\cite{tipslicingsurvey,whylineicse}.

These efforts have shown some of the promise of provenance, but are not a
debugging panacea.  First, prior work~\cite{badapples} has focused on providing
provenance information for only some unusable {\em values}, like {\tt NULL}.
Limiting the applicability of their technique to only some failures, like {\tt
Null Pointer Exceptions}.  Second, while prior work has shown that limited
provenance tracking for Java programs is possible with overheads low enough for
production~\cite{badapples}, a more general form of provenance tracking for C
and C++ programs has been elusive.  Prior approaches that use
slicing~\cite{thinslicing, tipslicingsurvey, whylineicse} or heavy-weight
memory checkers~\cite{badapples} impose overheads too high for production for C
and C++ programs. Third, no prior technique that we know of has collected
provenance information that is tailored to debugging concurrent programs with
overheads that are low enough for use in production systems.  Addressing
concurrency is important because a value could have originated in any of a
large number of threads, making reasoning about why a variable contains that
value complex and difficult.

In this work we design a novel form of in-production provenance tracking for
concurrent programs with three goals.  (1) To collect a general form of
provenance information, not limited to only certain values.  (2) To provide
provenance information that helps understand how threads interact, which is the
key to understanding concurrency bugs.  (3) To work with overheads low enough
for use in production (ideally, less than 10\%).

We develop a novel provenance tracking mechanism called {\em last writer
slicing} that satisfies these goals.   As the program executes in production,
last writer slicing tracks the thread and program point that last wrote any
value to each potentially shared memory location --- its provenance.  Last
writer slices are saved with a core dump when the program fails and are
available in the debugger ({\em e.g.}, for breakpoint debugging).

Last writer slices are primarily useful for concurrency debugging, but we also
observe that provenance information in the form of last writer slices is a
useful foundation for implementing other concurrent program analyses.  Several
tools from prior work has used {\em ad hoc} last-writer information as part of
more heavy-weight analyses in domains like bug detection ({\em e.g.},
~\cite{dmtracker,avio}), automatic debugging ({\em e.g.},
~\cite{recon,conseq,cci}), profiling ({\em
e.g.},~\cite{threadclustering, schedpredictionmodel}), and program
understanding~\cite{oshatr}.

Motivated by these prior {\em ad hoc} efforts, we build {\em Communication
Traps} or {\em \ctraps}  on top of last writer slices
(Section~\ref{sec:ctraps}).  \ctraps checks the last writer slice of each
potentially shared memory location when a thread accesses such a location.  If
the accessing thread is different from the thread that last wrote the location,
the access signifies {\em communication} between the threads.  When
communication occurs, \ctraps delivers a software {\em trap} to the accessing
thread, giving it information about how communication occurred.  The trap can be
handled by an analysis-specific trap handler.  Trap handlers implement {\em
\ctraps applications} through an extensible interface that can interpose on the
execution (Section~\ref{sec:apps}).  By default, \ctraps has very low overheads
and similarly to last writer slicing are appropriate for use in production in
many cases.  \ctraps applications can perform arbitrary concurrency analyses,
but at their own discretion and cost.  As we show in our evaluation, the
overhead of \ctraps applications scales with the complexity of their analysis.

%Provenance is also useful in implementing concurrent program analyses.
%Dependence information is also an important part of many program analysis
%tools.    Bug detection~\cite{avio,fasttrack,raceslicing,dmtracker} and
%debugging~\cite{tipslicingsurvey,bugaboo,recon,cci,defuse,conseq,falcon} tools
%analyze inter-thread dependences to find buggy behavior and help programmers
%understand their errors.
%Profilers~\cite{threadclustering,schedpredictionmodel} analyze inter-thread
%dependences to find optimization opportunities.  Dependences are also essential
%to some program understanding systems~\cite{oshatr} and dynamic checkers for
%thread interaction properties~\cite{oshajava,velodrome}.
%Record-Replay~\cite{chimera,doubleplay,fdr,rtr} systems track and record the
%order of dependent events to help reproduce failures.  These systems would be
%even more valuable, if their underlying dependence analysis was run in
%production.  However, the heavy-weight analyses proposed by many of these
%systems impose overheads too high for production use (5-100x
%slowdown)~\cite{raceslicing,dmtracker,velodrome,recon,defuse,oshajava,chimera,coredet,stm}.
%

%First, we describe {\em last writer slices}
%(Section~\ref{sec:lastwriterslices}).       Last writer slices are
%useful because they expose the provenance of a memory location's contents
%during debugging (Section~\ref{sec:debugging}).  When a failure occurs, last
%writer slices are saved with a core dump.  Last writer slices are also
%available during an execution in the debugger ({\em e.g.}, at breakpoints).
%Note that while prior work ({\em e.g.},
%~\cite{recon,fasttrack,conmem,conseq,cci}) has relied on {\em ad hoc}
%last-writer analyses, in this work, we isolate the essential last-writer
%information in the form of last writer slices and provide an implementation
%with overheads appropriate for production systems.  Tracking last writer slices
%has low overhead because it imposes overhead on only write operations:  reads
%are overhead-free.  In Section~\ref{sec:eval} we experimentally show that
%collecting last writer slices indeed has overheads suitable for production
%systems for a variety of applications.

%\cite{defuse,conseq,recon,bugaboo,raceslicing,fasttrack,falcon}
%\cite{oshajava,osha} 
%\cite{avio,dmtracker,cci,daikon}
%\cite{aviso,cfix} 
%\cite{mtperf,criticalitypredictors,schedpredictionmodel}  
%\cite{defuse}
%\cite{avio,cci,defuse,recon,bugaboo,falcon,dmtracker,aviso,cfix,criticalitypredictors,schedpredictionmodel}

%Deterministic execution
%systems~\cite{coredet,grace} monitor dependences to determine how and when to
%serialize portions of threads' executions to enforce determinism.  Software
%trasactional memory (STM) systems~\cite{stm} monitor dependences to identify
%conflicts between concurrently executing atomic code regions.

%Our goal in this work is to collect, with deployable efficiency, useful
%dependence information that captures data provenance for use in debugging and
%serves as a foundation for more sophisticated program analyses.  We start from
%two simple observations: (1) Reads are frequent so we must minimize tracking
%overheads on reads, and (2) In concurrency analyses, {\em inter-thread}
%dependences are most important.  Based on these observations, we focus first on
%tracking write-after-write dependences (WAWs), which provide a useful form of
%data provenance information.  Building on that, we describe support for
%tracking {\em communication dependences}.  Communication occurs when a thread
%reads or overwrites a memory location that was most recently written by a
%different thread; {\em i.e.}, inter-thread read-after-write dependences (RAWs)
%as well as WAWs.  In this work, we demonstrate that this form of light-weight
%dependence tracking is useful and versatile, and has inherent overheads that
%are low enough for use in deployed systems.  There are two parts to our work:

\paragraph{Contributions of This Work}
To summarize, this work makes several contributions:
\begin{itemize}

\item{We describe last writer slicing, an novel data provenance tracking
mechanism, that produces information that is useful for debugging concurrent
programs.  We design and implement system support for collecting last writer
slices that is efficient enough for use in production.  }

\item{Building on last writer slicing, we develop \ctraps, an extensible
framework for implementing analyses of inter-thread communication in
multi-threaded programs.  }

\item{We evaluate the utility of last writer slices and \ctraps. We first
present a study of several real concurrent program bugs and show that last
writer slices are useful for debugging.  We then show how \ctraps is useful for
building dependence analyses by implementing two analyses from prior
work~\cite{cci,defuse,recon}. }

\item{We evaluate our system's performance using a variety of important
applications ({\em e.g.}, MySQL, memcached and the PARSEC
benchmarks~\cite{parsec}).  We show that in many cases our designs have
overheads low enough for use in deployment (0--15\%).}

\end{itemize}

%In this work, we begin by showing that the cost of monitoring communication has
%been overstated in prior work.  We do so by designing support to monitor
%communication that, for a variety of important applications ({\em
%e.g.},databases, key-value stores, web servers) the overheads are extremely low
%-- so low that communication tracking is viable even for today's deployed
%systems.  We illustrate our claim with system support for  {\em Communication
%Traps} or \ctraps.  Our support for \ctraps tracks communication between
%threads and exposes communication events to trap handlers that can implement
%applications, such as the above.  We show the feasibility of our approach by
%implementing a selection of the above applications, as well as several novel
%applications.  We study the performance characteristics of \ctraps and provide
%a comparison to prior work, showing that our system has lower performance
%overheads -- in some cases, by one to two orders of magnitude lower.  We then
%show that even in cases that, for \ctraps, incur higher performance overheads,
%judicious use of sampling reduces overheads to an acceptable level, and retains
%many of the benefits of non-sampling \ctraps.

\section{Background}

In this section we discuss debugging with data dependence information, using an
example, in order to motivate collecting last writer slices in deployed
systems.  We then provide context for \ctraps by discussing some more
sophisticated concurrency analyses that rely on the kind of dependence
information we are focusing on in this work.

%Many systems that support programming and execution models, debugging tools,
%and other software engineering tools track dynamic memory dependences during a
%program's execution.

\subsection{Debugging and Data Provenance}

Data provenance information is directly useful during program debugging.  prior
work~\cite{whylinechi,whylineicse} showed that when programmers debug
programs, they benefit from the ability to directly ask such provenance
(``why'') questions of their debugger.  Origin tracking~\cite{badapples} and
program slicing~\cite{tipslicingsurvey} also make debugging easier by reporting
the provenance of values.  Like these prior efforts, in this work we focus on
using data provenance information for {\em debugging} -- understanding the root
cause of a failure -- as opposed to {\em bug detection} -- finding new, unknown
bugs.

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{figs/JDKStringBufferFail.pdf}
\caption{\label{fig:coreDumpFail}{\bf Atomicity violation bug in a string buffer implementation.} The core dump does not reveal the connection between {\tt append()} and {\tt erase()}, which is crucial to understand the bug.}
\end{figure}

Figure~\ref{fig:coreDumpFail} illustrates the role of provenance information in
debugging, using buggy code from the JDK-1.4 stringbuffer library, adapted to
C++ by other researchers and used variously in prior
work~\cite{concurrencybugs}.  The stringbuffer object pointed to by {\tt sb} in
Thread 1's code contains a buffer and a variable called {\tt count} that stores
its length.  The {\tt append()} code should atomically read {\tt sb->count} and
update the string's buffer (the full update code is not pictured).  The failure
occurs when the {\tt append()} code in Thread 1 is interleaved with the {\tt
erase()} code in Thread 2 as indicated in the figure.  The {\tt erase()} code
updates the buffer's {\tt count}, making the version that Thread 1 read into
{\tt len} stale.  Thread 1's assertion fails at {\bf B} using the stale {\tt
len} value, causing a crash.

When the failure manifests, Thread 1 is at the assertion in {\tt append()}, as
pictured.  However, Thread 2 may have completed its erase operation and moved on
to execute some unrelated code.  The call stack for Thread 2 may show nothing
about its involvement in the failure, despite the fact that its interleaving
with the {\tt append()} code is the root cause.  Examining the core dump, the
programmer can see that the length and buffer are inconsistent, but is left to
ask {\em why} they are inconsistent.

Understanding the connection between {\tt append()} and {\tt erase()} is the
key to fixing this bug.  The provenance of the value stored in {\tt sb->count}
at {\bf B} was the write at {\bf A} in {\tt erase()}.  Provenance information
shows the programmer that key connection.  Prior provenance work on origin
tracking for unusable values~\cite{badapples} does not help here.  The involved
values are inconsistent, but they are all {\em usable} so their provenance
would not be tracked.

Another challenge posed by this bug is that it manifests as a failure
infrequently and under only real-world conditions.  Prior slicing and
provenance debugging
techniques~\cite{tipslicingsurvey,thinslicing,whylineicse}, have overheads too
high to use in production.  Used offline, these techniques may not observe the
failure in a reasonable amount of time and triggering the bug may require
unavailable real-world inputs making offline analysis impossible.   A better
strategy is to continuously collect that information in deployed systems and
package it with core dumps sent to programmers with bug reports.  In this work,
we propose a last writer slicing design that espouses this {\em in situ} data
collection strategy and does so with overheads appropriate for production
systems.

\subsection{Communication Tracking in Concurrency Analyses}
\label{sec:background:comm}

Many useful concurrency analyses rely on monitoring inter-thread
communication.  The domains of these analyses include 
debugging~\cite{defuse,conseq,recon,bugaboo,raceslicing,fasttrack,falcon},
software engineering~\cite{oshajava,oshatr}, anomaly
detection~\cite{avio,dmtracker,cci,daikon}, bug and failure
avoidance~\cite{aviso,cfix} and concurrent performance
profiling~\cite{threadcriticality,schedpredictionmodel}.  The breadth of
applications illustrates the generality of communication tracking.

Despite the potential of a general communication tracking mechanism, prior work
has largely relied on {\em ad hoc} communication analyses.  For example,
DefUse~\cite{defuse} tracks communication via read-after-write (RAW)
dependences, but not write-after-write (WAW) dependences.  Recon~\cite{recon}
tracks some inter-thread WAW and RAW dependences.  These analyses could both be
built with a general communication tracking mechanism.  Unfortunately, their
single-purpose implementations track slightly different information, making
them incompatible and difficult to compose.

These analyses vary in their purpose and environments, but a cross-cutting
theme is the advantage reaped from their use in a deployment environment.
Debugging tools, profilers, and anomaly detectors benefit from seeing diverse,
real-world behavior.  Dynamic failure avoidance techniques must work in
deployed systems to be effective.  Deployment environments demand low overheads
and all of these analyses benefit from a general, high-performance
communication tracking framework.  Unfortunately, many of these techniques are implemented using heavy-weight infrastructure, like binary instrumentation~\cite{pin}, making them inappropriate for use in production.

There is a need for a general communication tracking framework that has
performance overheads that are low enough to use in production systems.  As we
describe in Section~\ref{sec:ctraps}, we aim to satisfy that need in this work.

%For example, in a software transactional memory (STM) system~\cite{stm}
%communication corresponds to interference between accesses in different threads
%that must be detected to rollback conflicting transactions.  Deterministic
%concurrent runtime systems~\cite{coredet,grace} similarly monitor
%communication, to apply deterministic ordering constraints.  STMs and
%deterministic systems must track all dependences and they must track them
%precisely, to provide strong guarantees as programming and execution models.
%Unfortunately, precision has a high runtime overhead -- {\em e.g.}, CoreDet
%incurs a mean overhead of around 5x~\cite{coredet}.  This overhead is too high
%for use in production.

\section{Last Writer Slices}
\label{sec:lastwriterslices}
A last writer slice is a dynamic property of a memory location at a point in a
program's execution.   A memory location's last writer slice is a record
containing the program point and the identifier of the thread that last wrote
data to that memory location.

We gather last writer slices at runtime by maintaining a {\em last writer
table} (\lwt) that maps from a memory location's address to its last writer
slice.  Just before a thread writes to a memory location, the thread first
finds the memory location's entry in the \lwt and updates that entry with its
thread identifier and its current point in the program.  After updating the
\lwt, the thread performs the original memory access.

Figure~\ref{fig:basicLWT} shows the basic operation of the \lwt.  The right
side of the figure shows a program execution with three threads.  On the left
is the \lwt at the point of each memory access.  There are two key things to
notice.  First, write operations update the \lwt entry of the memory location
they access.  Second, there is no action necessary for read operations.
Making read operations cheap is key to making the overhead of last writer
slicing low enough for deployment use.

{\bf TODO: Fix this figure to reflect availability of LWS info if the program crashes/is stopped at *any* point in the exec (like the read X in thread 2).}

\begin{figure}[h]
\centering
\includegraphics[scale=.6]{figs/BasicLWT.pdf}
\caption{\label{fig:basicLWT}Basic operation of the last writer table. Different shapes signify different distinct memory operations. }
\end{figure}

\subsection{Debugging with Last Writer Slices}
\label{sec:debugging}
%To answer data provenance questions during debugging, prior work has mostly
%used dependence slices that include intra-thread
%dependences~\cite{whylineicse,conseq,tipslicingsurvey}.  The overhead of
%slicing dependence chains and tracking temporal sequences is often high,
%relegating such techniques to working on traces~\cite{whylinechi} or debugging
%executions run ``in the lab''.

At each point in an execution, a memory location's \lwt entry represents the
provenance of the data stored in that memory location.   Last writer slicing
performs no action on read operations and to be useful, it does not need to.
Instead, a programmer can examine an \lwt entry at any arbitrary point in an
execution ({\em e.g.}, at a breakpoint), including after a failure, and see a
value's provenance.  Including the \lwt with core dumps means provenance is
available when post-mortem debugging production failures.

Figure~\ref{fig:jdklws} illustrates how the information in the \lwt helps debug
the JDK stringbuffer bug introduced in Figure~\ref{fig:coreDumpFail}.  Recall
that the failure occurs because Thread 2's execution of {\tt erase()} violates
the atomicity of Thread 1's execution of {\tt append()}.  The \lwt entry for
{\tt sb->count} at {\bf B} shows the programmer that its last writer was Thread
2 at {\bf A}. Thread 2 may have run ahead to an arbitrary point in its
execution.  Nothing in the core dump implicates {\tt erase()} in causing
this bug.  The last writer slice reveals the interference between {\tt append()}
and {\tt erase()} via {\tt sb->count}, helping the programmer understand the
bug's cause.

%The core dump does not provide any
%insight into why the mutex is invalid at the time of the call or what its
%value should have been instead.

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{figs/JDKStringBufferDebug2.pdf}
\caption{\label{fig:jdklws}{\bf Using last writer slices to debug an
ordering violation in the JDK 1.4 StringBuffer library.} The last writer slice reveals the interaction between the erase function and the append function that leads to a failure.}
\end{figure}

Another important trait of last writer slices illustrated by
Figure~\ref{fig:jdklws} is that when the assertion fails at {\bf B}, stopping
the execution, the \lwt is saved and can be sent to the programmer with the
core dump.  Capturing good debugging information is especially important for
concurrency bugs that manifest very rarely or in the presence of only peculiar
real-world inputs.  A primary goal of this work is to show that we can collect
last writer slices with overheads that are low enough for production use so
they can be included pervasively in bug reports.

%We discuss last writer slices in more depth by first describing our run time
%support for collecting them.  The system support consists of a runtime library
%and a compiler analysis.  Section~\ref{sec:debugging} shows how last writer
%slices help with debugging using our running example from
%Figure~\ref{fig:bgcoredumpfail}.  We formalize last writer slices in
%Section~\ref{sec:lwssoundness}.

\subsection{Runtime \& Compiler Support}

Figure~\ref{fig:lwsoverview} shows an overview of our system support for last
writer slicing.  The programmer writes their program as usual. Our compiler
instruments write operations in the program with calls to our runtime that
update the \lwt.  The \lwt is maintained in the runtime and saved for use
during debugging, like the core dump.

\begin{figure}[h]
\centering
\includegraphics[width=.9\columnwidth]{figs/LWS_Overview.pdf}
\caption{\label{fig:lwsoverview}{\bf Overview of system support for last writer slicing.} Our compiler and runtime library ensure the \lwt remains updated so it can be used during debugging.}
\end{figure}

The \lwt implemented in the runtime library is a {\em shadow memory} that
holds one entry for each potentially shared memory address accessed by the
program.  Each \lwt entry records the thread and program point that last wrote
to the entry's corresponding memory location.  When a thread performs a write
to a memory location, it updates the \lwt by recording its thread identifier
and the program point of its write operation to the memory location's \lwt
entry.  A program point may be an instruction address, or a call stack.  We use
instructions by default, but we experimented with 2-address bounded call-stacks
as well.

Collecting last writer slices requires instrumentation on each write operation.
We built an instrumenting compiler pass that inserts a call to an \lwt update
function in our runtime library just before each write that may access data
last written by another thread (escaping memory accesses).  Write
instrumentation calls are passed three pieces of information: the address of
the memory word involved in the access, the thread identifier of the accessing
thread, and the accessing program point.

We use a {\tt GCC}'s local function escape analysis to identify accesses to
memory locations that are provably inaccessible to code running in other
threads.  Such operations access unshared data and they are not instrumented by
our compiler.  
%Using an inter-procedural thread escape analysis may eliminate more
%instrumentation sites.

\subsection{\lwt Formalism and Correctness}
{\bf TODO: Necessary?  Is it really interesting?}
\label{sec:lwssoundness}
To discuss the correctness of our \lwt-based last writer slice support, we
first develop a formal description of last writer slices.  We consider an
N-threaded program $P = \{P_1, P_2, \ldots P_N\}$, where $P_t = (p^{t}_{1},
p^{t}_{2}, \ldots, p^{t}_{k_{t}})$ is the sequence of $k_{t}$ dynamic
instructions executed by thread $t$.  Without loss of generality, we assume all
instructions are memory reads and writes, with the form $Op(m)$ where $Op$ is
$Read$ or $Write$ and $m$ is a uniquely addressed memory location.
A point in an execution of $P$ after $e$ instructions have
executed is defined by $Q = \{q_{0}, \ldots, q_{e}\}$ where each $q \in Q$ is a
$p^{t}_{j}$ from one of the $k_t$ instructions of thread $t$.  We introduce
$LWS[m]$ -- the Last Writer Slice for memory location $m$.   At any point $Q$,
$LWS[m]$ holds $(t,p^{t}_{\ell w})$, where $p^{t}_{\ell w} = max_{\ell}( 
q_{\ell} = Write(m) \in Q)$ and $t$ is the thread that
executed $p^{t}_{\ell w}$. Informally, $p^{t}_{\ell w}$ is the last instruction
in $Q$ to write $m$ and $t$ is the thread that executed it.  For a
never-written location $n$, $LWS[n] = (\emptyset,\emptyset)$ indicating it is
unwritten.

For our design to be correct, $LWT[m]$, the entry for a memory location $m$ in
our \lwt implementation, must be the same as $LWS[m]$ at all points in an
execution.  Correctness in the \lwt requires {\em soundness} and {\em
completeness}.  A {\em sound} design only updates $LWT[m]$ when $m$ is written.
A {\em complete} design always updates $LWT[m]$ when $m$ is written.  Our
compiler inserts \lwt updates before every write operation (for completeness)
and only before write operations (for soundness).  Formally, given a program,
$P$, our compiler pass transforms it to a new program, $P'$.  Each thread in
$P'$ executes a modified sequence of instructions: $P'_{t} \in P' = \{
I(p^{t}_{1}), p^{t}_{1}, \ldots, I(p^{t}_{k_{t}}), p^{t}_{k_{t}} \}$.
$I(p^{t}_{j})$ sets $LWT[m] = (t,p^{t}_{j})$ if and only if $p^{t}_{j} =
Write(m)$ and is a no-op otherwise.  This reasoning shows our \lwt is both
sound, complete, and thus correct.

Next we discuss two nuances of our design: (1) the interaction of our \lwt
implementation with code containing data-races; and (2) the correctness of the
\lwt when a program addresses data that overlap in memory.

\paragraph{Consistency and Atomicity}

Our implementation of the \lwt does not add any synchronization to the program.
As a result our analysis does not incur the overheads or possible thread
schedule perturbation synchronization can cause.  However, there are two
potential issues that stem from this design choice. First, we do not enforce
memory consistency of meta-data reads and writes in different threads.  Second,
our implementation does not enforce the atomicity of \lwt updates and their
corresponding program accesses.  In the absence of data-races our \lwt
implementation is {\em always correct} and neither of these issues exist.  Both
atomicity and ordering is enforced by the program synchronization that ensures
consistency for the data being written by the instructions our system
instruments.

We use our \lwt formalism to show that without data-races, \lwt updates and
their corresponding memory accesses are atomic and well-ordered.  The relation
$p \prec q$ denotes that $p$ happens before $q$.  In correctly synchronized
code, any write by thread $t$ to $m$, $p^{t}_{j} = Write(m)$, is ordered by
program synchronization with $p^{t'}_{j'} = Op(m)$, a later access to $m$ by
any other thread $t'$ -- {\em i.e.}, $p^{t}_{j} \prec p^{t'}_{j'}$.  Our
compiler inserts an \lwt update $I(p^{t}_{j})$ into $P'$ before its write
operation, $p^{t}_{j}$. Thus, $I(p^{t}_{j}) \prec p^{t}_{j}$ due to program
order.  Happens-before is transitive, so without data-races, $I(p^{t}_{j})
\prec p^{t}_{j} \prec I(p^{t'}_{j'}) \prec p^{t'}_{j'}$, meaning $I(p^{t}_{j})$
happens before all accesses that $p^{t}_{j}$ happens before {\em and before
their respective \lwt updates}, $I(p^{t'}_{j'})$.  Symmetrically, for any
access $p^{t''}_{j''}$ in another thread $t''$ that precedes $p^{t}_{j}$,
$p^{t''}_{j''} \prec p^{t}_{j}$.  Our compiler inserts each $I(p^{t}_{j})$
immediately before each $p^{t}_{j}$, so $I(p^{t''}_{j''}) \prec p^{t''}_{j''}
\prec I(p^{t}_{j}) \prec p^{t}_{j}$, happens-before ordering $I(p^{t}_{j})$
after all preceding accesses to $m$ in other threads, {\em and after their
repective \lwt updates}.  Thus, at all points $Q$ in a data-race free
execution, given $q_r = I(p^{t}_{j})$, $q_p = p^{t}_{j}$, and $q_r \prec q_p$,
$\forall_{ \{q_s | q_r \prec q_s \prec q_p \}}: q_s \ne Op(m) \wedge q_s \ne
I(Op(m)) $.  In other words, \lwt updates and their accesses are atomic.

In a program with a data-race, there are some accesses, $p^{t}_{j}$ and
$p^{t'}_{j'}$, that are concurrent.  The \lwt updates for these accesses are
also concurrent because they are not transitively ordered by the accesses' ordering.  The unordered accesses and \lwt updates can interleave arbitrarily
because they are concurrent.  Some racy thread schedules may violate the
atomicity of accesses and their \lwt updates.  Such an execution could reach a
state $Q_{bad} = (\ldots, I(p^{t}_{j}), I(p^{t'}_{j'}), p^{t'}_{j'}, p^{t}_{j},
\ldots)$ and in that state, $LWS[m] = (t,p^{t}_{j})$, but $LWT[m] =
(t',p^{t'}_{j'})$.  The \lwt was effectively not updated when $p^{t}_{j}$
executed (violating completeness), or was effectively updated to reflect that
$p^{t'}_{j'}$ was $m$'s last writer, when $p^{t}_{j}$ was (violating
soundness).

According to the C++11 language specification, in the presence of a data-race,
a program's behavior is undefined and arbitrary state corruption can occur.  In
the presence of data-races, {\em any} analysis, including our \lwt analysis,
running on such a program could experience arbitrary corruption.  In practice
({\em e.g.}, on x86) these atomicity violations may cause limited \lwt
corruption, but even this is unlikely.  These atomicity violations only occur
only when accesses to the same data occur within a very short time interval,
which is uncommon, {\em and} an unserializable interleaving of accesses and
\lwt updates also occurs.

%{\bf TODO: Figure?}

\paragraph{Non-overlapping Memory References} 
We assume each written memory location corresponds to a non-overlapping memory
region because our \lwt design is {\em granularity oblivious}.  We discuss this
issue informally due to space constraints.  Each \lwt entry corresponds to a
memory location referenced by a program instruction.  Two instructions might
reference different locations that overlap because the referenced locations
have different granularities ({\em e.g.}, accessing a byte that is part of a
previously accessed word).  When that happens, both accesses update different
\lwt entries, because they reference different locations.  However, the
overlapping region then has two different entries reflecting the last {\em two}
updates to its parts rather than just its last write.

The \lwt remains correct, assuming the program only accesses non-overlapping
regions.  In programs that access overlapping regions, there is always some
entry in the \lwt that correctly reflects each memory location's last writer.
However, the entry for a coarse-grained region may be obsolete because of an
access to an overlapping sub-region.

One approach to mitigating this issue would be to allocate \lwt entries for
regions of fixed granularity.  We chose not to do so for two reasons.  First,
associating \lwt entries with regions of  large granularity only ({\em i.e.},
cache-line) precludes sound analysis for fine-grained accesses.  Second,
associating \lwt entries with regions of small granularity ({\em i.e.}, byte)
requires multiple \lwt updates for larger granularity accesses, increasing
their overheads proportionately to the granularity of the access.

\section{Communication Traps}
\label{sec:ctraps}

\ctraps is a framework for analyzing and interposing on inter-thread
communication events during a program's execution.  We build \ctraps on top of
last writer slicing that we described in Section~\ref{sec:lastwriterslices}.
Using last writer slice information, \ctraps dynamically tracks 
inter-thread communication events. \ctraps exposes communication
events as {\em traps} during an execution that {\em trap handlers} can handle
to perform analysis and interposition.  We formalize trap delivery and handling
in Section~\ref{sec:ctsoundness}.

\subsection{\ctraps Design}

We now discuss \ctraps in more detail by describing the design of our \ctraps
system support.  Figure~\ref{fig:systemdiagram} shows an overview of our system
support.  The programmer writes their program as usual.  They compile it using
the \ctraps compiler. The \ctraps compiler includes the compiler support for
collecting last writer slices.  In addition, the \ctraps compiler also inserts
calls to the \ctraps runtime at points where communication may occur.  The
resulting compiled executable links to the \ctraps runtime that collects last
writer slices and monitors communication.  The runtime loads and manages \ctraps
handlers when the execution starts via a plugin interface.  \ctraps handlers,
written independently of the original program implement \ctraps applications.
Traps are delivered to handlers by the runtime when operations communicate.
Together, the \ctraps executable and the \ctraps handlers are deployed.

\begin{figure}[htb]
\centering
\includegraphics[width=.90\columnwidth]{figs/CTraps_Overview.pdf}
\caption{\label{fig:systemdiagram}{\bf CTraps system support.}}
\end{figure}

%Our system support for \ctraps incorporates the support for collecting last
%writer slices described in Section~\ref{sec:lastwriterslices}.  In addition,
%our \ctraps system support also monitors all read and write operations that
%might communicate and traps on all those that communicate.

\subsubsection{Compiler and Runtime Support}

\ctraps delivers software traps before all read and write operations executed
by one thread that access a memory location that was last written by another
thread, {\em i.e.}, communicating memory operations.  Traps are simply function
calls made before such operations execution.  \ctraps implements traps using a
combination of compiler and runtime library support.  Note that we use the name
``trap'' for the interposition points in \ctraps, but we implement traps using 
function calls only, as opposed to signals or some other mechanism.

\paragraph{Compiler Support}
Our compiler inserts a call to the \ctraps {\em trap hook} function immediately
before all read and write operations that access potentially escaping memory
locations.  Trap hooks are passed the address of the memory location being
accessed, the thread identifier of the accessing thread, the type of operation
being performed, and the program point of the access.

\paragraph{Runtime Support}
The \ctraps runtime implements the trap hook function,  which has two purposes.
The trap hook function detects communication and delivers traps to \ctraps
applications when communication occurs.  The trap hook function detects
communication by looking up the \lwt entry for the memory location being
accessed.  If the thread identifier in the \lwt entry is different from the
accessing thread, then the trap hook function delivers a trap to the accessing
thread, just before the memory access is allowed to proceed.  To deliver a
trap, the trap hook function calls {\em trap handler routines} that are defined
in \ctraps applications.  Trap handler routines are like signal handlers.  They
are registered with the \ctraps runtime when the program starts.  Trap handlers
may contain arbitrary application-specific code.  Figure~\ref{fig:hookapi}
shows the \ctraps trap handling API.  Trap handlers are passed the four pieces
of information passed to the trap hook, as well as the \lwt's record of the
program point and thread identifier of the last write to the involved memory
location.

\begin{figure}[htb]
\centering
\begin{verbatim}
void trap(AccessType type, 
          ThdId currThd, ProgPt currPt,
          ThdId lastThd, ProgPt lastPt);
\end{verbatim}
\caption{\label{fig:hookapi}The \ctraps trap handling interface.}
\end{figure}

%MAYBE PUT THIS BACK IN
%\subsection{Formalism and Correctness}
%\label{sec:ctsoundness}
%To discuss the correctness of \ctraps, we first formalize trap delivery by
%building on our last writer slicing formalism from
%Section~\ref{sec:lastwriterslices}.  
%
%Given a program $P'$ already instrumented with \lwt updates, the \ctraps
%compiler produces $P''$. A thread $t$ in $P''$ executes a modified instruction
%sequence, $P''_{t} = \{ T(p^{t}_{j}), I(p^{t}_{1}), p^{t}_{1},$ $\ldots,
%T(p^{t}_{j}), I(p^{t}_{k_{t}}), p^{t}_{k_{t}} \}$, where $T(p^{t}_{j})$ is a
%trap hook call.  We consider a point $Q = \{q_1, \ldots, q_{e-2}\}$ after $e-2$
%instructions of an execution have executed and $q_{e} = p^{t}_{j}$ is about to
%execute after its trap hook $T(p^{t}_{j})$ and \lwt update $I(p^{t}_{j})$
%execute.  The trap hook $T(p^{t}_{j})$ {\em delivers} a trap before if $LWT[m]
%= (t',p^{t'}_{\ell w})$ and $t \ne t'$.  Informally, the trap hook delivers a
%trap before the instruction if the thread that last wrote $m$ is different from
%the one executing the instruction.  When a trap hook, $T(p^{t}_{j})$, delivers
%a trap, thread $t$ executes an arbitrary trap handler $H = \{h^{t}_{1}, \ldots,
%h^{t}_{h_{n}}\}$.  Thus, after the trap, the \lwt update, and the instruction
%execute, the execution reaches the new point $Q = (q_{1}, \ldots, T(q_{e}),
%h^{t}_{1}, \ldots, h^{t}_{h_{n}}, I(q_{e}), q_{e})$, and continues.
%
%A \ctraps implementation must be sound and complete to be correct.  A design is
%{\em sound} if it only delivers a trap when threads have communicated.  A
%design is {\em complete} if it always delivers a trap when threads have
%communicated.  The \ctraps compiler puts a trap hook on all read and write
%operations, and only on those operations.  Assuming a correct \lwt, \ctraps is
%correct.  Our \ctraps implementation also does not insert any synchronization
%into the program, so it comes with the same caveats related to data-race
%freedom as the \lwt.  The same arguments for \lwt correctness apply to the
%correctness of trap delivery and handling.  Most importantly, data-race freedom
%implies that trap hooks, \lwt updates, and accesses are all atomic.

%\paragraph{Escape Analysis}
%All our designs eliminate instrumentation on accesses to memory locations
%statically guaranteed to be inaccessible to other threads.  Doing so has no
%effect on soundness or completeness because accesses to such locations are
%guaranteed never to be communicating.

\subsection{\ctraps Applications}
\label{sec:apps}
\ctraps applications are implemented as shared library plugins that are loaded
by our runtime.  They must implement the \ctraps trap handler API, which
requires a trap handler function and permits a constructor, destructor, thread
constructor, and thread destructor, which allow initialization and disposal of
global and thread-local state.  Many useful \ctraps applications are possible,
several of which were outlined in Section~\ref{sec:background:comm}.    To
demonstrate that \ctraps can implement real analyses with little work, we
implemented two applications. The first application is a version of a debugging
analysis from CCI~\cite{cci}.  The second application is a communication graph
collector, similar to DefUse~\cite{defuse}, Recon~\cite{recon}, and
DMTracker~\cite{dmtracker}.

\paragraph{CCI-Prev Implementation}
CCI-Prev is a technique from CCI~\cite{cci} that records a set of code points
that access a memory location when the previous access to the same location was
by a different thread.  We implemented a variant of CCI-Prev using \ctraps that
records such a set of code points.  In our implementation, communcating read
operations update the LWT like writes normally do.  Under this policy, any
operation that accesses a location that was last accessed by a different thread
is recorded.  The set of recorded accesses is the output of the analysis.  Our
\ctraps implementation of CCI-Prev took about 10 lines of code (on top of our
base system).

\paragraph{Communication Graph Implementation}
Communication graphs are the basis of several prior debugging
techniques~\cite{recon, bugaboo, defuse}.  We implement communication graph
collection using \ctraps.  When a trap is delivered, our implementation records
communication graph edges composed of the code point of the last writer to the
location being accessed and the code point of the trapping access.  The set of
recorded edges is the tool's output.  Our \ctraps implementation of
communication graphs took about 50 lines of code (including output formatting
code).

\paragraph{Other Applications}
In addition to those applications, we envision that a variety of applications
will be enabled by our simple interface.  Anomaly detection and ``likely
invariant'' techniques similar to AVIO~\cite{avio}, Daikon~\cite{daikon}, and
others are a natural fit for \ctraps.  \ctraps provides the support necessary
for implementing specification checkers for specifications related to
inter-thread communication~\cite{velodrome,oshajava}.  Guided testing
tools~\cite{cuzz,chess} and bug detection tools~\cite{ctrigger} focused on
concurrency could also be implemented as \ctraps applications.  \ctraps also
provides an interesting platform for developing new approaches to
record/replay~\cite{chimera,fdr}.  \ctraps is an enabling platform for
developing these or other analyses with low baseline overheads and without the
need to build an entire concurrency analysis from scratch.

\section{System Implementation}

We implemented last writer slicing and \ctraps.  We built our compiler support
as a plugin for GCC ({\tt gcc-4.7.0 20111112}).  We used GCC's points-to and
escape analysis support to prune instrumentation to non-escaping memory
locations.  Our compiler instruments calls to {\tt free} and {\tt delete} as
writes to the pointer being deallocated.  Our analysis handles all code
compilable using this version of GCC except for a few cases: we do not handle
accesses that compile to GCC {\tt BITFIELD\_REFERENCE} IR types because these
are undocumented, we do not handle inlined assembly instructions, and we do not
handle C++ exceptions.  Note that these limitations of our research prototype
are not fundamental and the limited cases are uncommon.

We built our runtime system from scratch.  We implemented the LWT as a fixed
size 2GB array of 64-bit words, fixing our memory overheads.  Each entry packs a program point and a thread
identifier into the 64-bit word.  By default, program points are single
instruction addresses, but we also support limited (2-address)
context-sensitivity by packing the current instruction's address and the
nearest return address into a 64-bit LWT entry. When an access to a memory
location occurs, we index into the LWT with the lower bits of the location's
address.  Our prototype implementation uses a lossy resolution policy for hash
collisions.  We use this policy in our prototype because it is faster and
simpler than a chaining concurrent hash table implementation.  Collisions are
extremely rare, so this simplification is unlikely to be a problem even in a
more industrial strength implementation.

\subsection{Release}
We released our implementation, free and open-sourced, on the
web~\cite{withheldforanonymity} and published it to the GCC plugin repository.
Since its release, we have seen contributions to our codebase from members of
the GCC development community and downloads from interested developers.

\section{Evaluation}
\label{sec:eval}
We evaluated last writer slices and \ctraps along several axes.  We show that
last writer slices provide information that is useful during debugging in
Section~\ref{sec:eval:debugging} using case studies on several buggy programs.
We evaluate our performance overheads and show that last writer slices and
\ctraps are feasible for use in deployed systems for a majority of workloads we
studied.  We then characterize our instrumentation and discuss the sources of
our system's run time overhead.   We show that \ctraps is useful by
implementing two existing applications and analyzing their performance.

\subsection{Debugging with Last Writer Slices}
\label{sec:eval:debugging}

We illustrate the debugging benefits of last writer slices by using our system
to debug several real-world programs studied previously in the debugging
literature.  Table~\ref{tab:bugs} provides an overview of the bugs we studied
describing the program, bug type, and bug report identifier where applicable.
We then provide a summary of last writer slicing's debugging benefits.  To do
so, we mirror the evaluation strategy of bad value origin
tracking~\cite{badapples}, as it is the only other low-overhead provenance
tracking mechanism we know of.  We evaluate our system using their three
evaluation criteria - (1) effectiveness, (2) triviality, and (3) usefulness.
Our technique is effective if the last writer slice reports a code point that
wrote a value that subsequently caused a failure.  A failure is trivially
debuggable if its cause is made obvious by examining the core dump alone, {\em
i.e.}, a last writer slice is unnecessary.  A last writer slice is useful if it
is effective and the thread and code point in the slice actually help
understand why the failure occurred or how to fix it.   To draw contrast to bad
value origin tracking, we indicate whether their analysis could help debug each
bug, based on the description of their technique.

\begin{table}
\scriptsize
\centering
\begin{tabular}{l|ll|l|ccc|c}
{\bf App} & {\bf Ver. \#} & {\bf Bug \#} &{\bf Type} & {\bf Eff.?} & {\bf Triv.?} & {\bf Usef.?} & {\bf O.T.?}\\ \hline
mysql  &  4.0.12 &  791     & Atom.  &Yes  &No    & Yes   & No  \\%All values usable
pbzip2 &  0.9.4  &  N/A     & Order  &Yes  &No    & Yes   & Yes \\
httrack& 3.43.9  &  N/A     & Order  &Yes  &No    & Yes   & No  \\%Used w/o origin - no help
jdk1.4 &  1.70   &          & Atom.  &Yes  &No    & Yes   & No  \\%All values usable
apache1&  2.0.48 &  21285   & Atom.  &Yes  &No    & Yes   & No  \\%All values usable
apache2&  2.2.9  &  45605   & Atom.  &Yes  &No    & Maybe & ??  \\
\end{tabular}
\caption{{\bf Debugging with last writer slices.} We describe each bug and summarize the debugging benefit of last writer slices using the evaluation criteria from ~\cite{badapples}.  }
\normalsize
\end{table}

\paragraph{Apache}
%http://issues.apache.org/bugzilla/show_bug.cgi?id=21287
%concurrency-bugs
We conducted a case study using Apache httpd-2.0.48, which has a bug in its
script cache. The cache uses reference counting to determine when to deallocate
cached objects.  When a thread sees a reference count hit zero, it frees the
object.  The bug is that the check that the count is zero and the deallocation
are not atomic but it should be.  Multiple threads may see the zero reference
count and free the object more than once, leading to a crash.

We debugged this issue by examining the core dump from a crashing execution.
The core dump showed that the memory allocator aborted due to one thread
freeing already freed memory at {\tt mod\_mem\_cache.c}, line 287.  We examined
the last writer table entry for the object being freed at that point.  That
entry revealed that a different thread last wrote that object at the same point
in the code.  The information in the last writer table illustrates why the
double free bug is happening -- two threads end up in the object deallocation
routine for the same object, which should not happen.

\paragraph{HTTrack}
%{\bf TODO: Deal with this duplication from the copy/paste}
%Figure~\ref{fig:coreDumpFail} illustrates the role that provenance information
%plays in debugging, using an example of a real bug from HTTrack, a web caching
%tool.  The figure depicts two threads manipulating a data structure that
%contains a mutex\_lock.  The program has an ordering violation bug: the
%program crashes when Thread 2 tries to acquire the lock before Thread 1 has
%initialized it.  Assuming the run time environment supports it, a core dump
%delivered to the programmer after such a failure reveals very little about how
%to fix the bug.  The lock is uninitizalized, so the core dump may show that it
%holds {\tt NULL} or it may hold random bits.  Both scenarios provide no
%insight into the root cause of the failure, which is that the code permits a
%use of the lock before its initialization.     This revalation is of limited use as it does not rule out
%arbitrary data corruption as the bug's root cause.
%
%\begin{figure}[h]
%\centering
%\includegraphics[width=\columnwidth]{figs/LWSHTTDebug.pdf}
%\caption{\label{fig:httlws}{\bf Using last writer slices to debug an
%ordering violation in HTTrack.} The provenance information stored in the last
%writer slice for the mutex variable helps understand the bug's root cause.}
%\end{figure}
%
%
%The vital, missing piece to solving this bug is provenance information.
%Debugging would be easier if the core dump included information about the
%thread and code point that last wrote the lock.  The programmer would see that
%no thread had initialized the lock in the failing execution, but that Thread 1
%initialized the lock in non-failing executions.  This information guides the
%programmer to fix the bug by adding synchronization that ensures the correct
%operation ordering.

Figure~\ref{fig:httlws} shows how last writer slices help understand the root
cause of the concurrency bug from Figure~\ref{fig:coreDumpFail}.  The program
crashes at point {\bf B} when it calls {\tt hts\_mutex\_lock()} with an invalid
mutex.    The root cause of this bug is that Thread 1 fails to initialize the
mutex before Thread 2 uses it.   The core dump is of little debugging use, only
showing that, indeed, the lock is uninitialized.  Even if the programmer
compared the core dump from the failing run to the memory image from a correct
execution -- a helpful debugging strategy -- they would find only that in a
correct execution, the lock contains valid mutex state.

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{figs/LWSHTTDebug.pdf}
\caption{\label{fig:httlws}{\bf Using last writer slices to debug an
ordering violation in HTTrack.} The provenance information stored in the last
writer slice for the mutex variable helps understand the bug's root cause.}
\end{figure}

Looking at last writer slices make this root cause clear: in the failing
execution, the last writer slice indicates the mutex is uninitialized; in the
non-failing execution, the last writer slice indicates the mutex was
initialized by Thread 1 at point {\bf A}. While the core dump provides no
information about the connection between point {\bf A} and point {\bf B}, the
last writer slice reveals that the programmer should ensure {\bf B} follows
{\bf A} to prevent the failure.  Note that even augmenting the core dump with
bad value origin tracking~\cite{badapples} does not help with this bug.  In the
failing execution, origin tracking shows that the unusable lock is
uninitialized, but in the non-failing exeuction, origin tracking does not
monitor the {\em usable} lock value.  By contrast, last writer slicing tracks
that crucial information, helping debug this complex concurrency bug.

\paragraph{PBZip2}
We conducted a case study using a buggy version of PBZip2-0.9.1.    The bug is
a use-after-delete bug that leads to an assertion failure.  A worker thread
tries to acquire a mutex after it has been deleted by the main thread during
shutdown, causing the assertion to fail.

Last writer slices provide the key piece of information to debug this failure.
When the program fails, the core dump shows that the mutex is invalid because
it has been deleted.  The mutex's last writer slice shows that the last write
to the mutex before the worker tried to acquire it was when the main thread
deleted it.  The last writer slice reveals the connection between the the two
pieces of code involved in the bug: the code that acquires the mutex and the
code that deletes the mutex.  Note that this connection is not obvious when the
program crashes because the main thread may have executed arbitrarily far from
where it deleted the mutex before the worker thread crashes.

\paragraph{MySQL}

We used last writer slices to debug an atomicity violation in MySQL-4.0.12 that
is a critical security vulnerability.  Figure~\ref{fig:mysqllws} shows the
buggy code.  Thread 1 rotates the database's log file, marking it closed,
opening a new log file, then marking it open again.  Log rotation should be
atomic, but the lack of synchronization allows Thread 2 to run its insert
during log rotation.  The insert code checks if the log is marked closed. If
the log is marked closed, the insert is not logged, manifesting the failure.
We added an {\tt abort} statement to the non-logging case because this behavior
is a security vulnerability that the bug report describes as ``critical''.
Adding this {\tt abort} is reasonable, as the bug report describes the
condition in the {\tt else} block as being an error.

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{figs/MySQLDebug.pdf}
\caption{\label{fig:mysqllws}{\bf Using last writer slices to debug an
atomicity violation in MySQL 4.0.12.} The last writer slice reveals the code that closes the log before a database insertion that goes unlogged, which is a security problem.}
\end{figure}

We triggered the bug by running an insert query during a log rotation.  The bug
manifested, stopping execution at {\bf C}.  As described in the bug report, the
query went unlogged and examining the core dump revealed the log was marked
closed at point {\bf B} ({\em e.g.}, {\tt log\_type} was {\tt LOG\_CLOSED}).
However, the core dump did not reveal {\em why} the log was marked closed
there.  The last writer slice at {\bf C} for the {\tt log\_type} variable shows
that it was written last by Thread 1 at {\bf A}.  Without the last writer
slice, the relationship between the log rotation code and the insert code is
not apparent. The last writer slice reveals that the log rotation code
interacts with the insert code incorrectly via the {\tt log\_type} variable.

\paragraph{JDK-StringBuffer}

The running example in Figures~\ref{fig:coreDumpFail} and ~\ref{fig:httlws},
and the text in Section~\ref{sec:debugging} describes the benefit provided by
last writer slices in debugging an ordering violation in the JDK-1.4 stringbuffer library.

\subsection{Performance Evaluation: Setup and Benchmarks}

We conducted our evaluation on a machine running Linux 2.6.27-7, with a 2.27GHz
8-core Xeon E5520 processor with 2-way SMT ({\em i.e.}, 16 execution contexts)
and 10GB of memory.  We evaluated our system using PARSEC, a set of
architectural benchmark programs~\cite{parsec} and a collection of real server
and desktop applications.

\subsubsection{PARSEC}

We ran PARSEC programs with their {\tt native} input set, and with each
program's 8 thread configuration. We have omitted three of the PARSEC
benchmarks because our compiler pass does not handle C++ exceptions.

\subsubsection{Servers}

\paragraph{MySQL-5.1.65} MySQL is an industrial-strength database
server. To benchmark MySQL, we used {\tt SysBench OLTP} running with its
default 8 threaded configuration, measuring MySQL's performance to be the
throughput reported by {\tt SysBench}.

\paragraph{Apache-2.4.3} 
Apache-httpd is a popular web server.  To benchmark Apache, we used {\tt
ApacheBench} to request a static {\tt html} page 1,000,000 times using 8
threads, measuring Apache's performance to be the throughput reported by {\tt
ApacheBench}.

\paragraph{LevelDB-1.5.0}
LevelDB is a carefully tuned high-performance key-value store written by Google
developers. To benchmark LevelDB, we used the included {\tt db\_bench} utility,
running its ``Read while Writing'' test with 8 threads.  We measure LevelDB's
performance to be the operation throughput reported by {\tt db\_bench}.

\paragraph{Memcached-1.4.4}
Memcached is an in-memory key-value store frequently used as a cache for web
services.  We were unable to find a standard benchmark for Memcached.  Instead,
we wrote a C program that uses libmemcached-0.49 to issue a
mixture of 10\% load requests and 90\% store requests for a single key 
simultaneously from 8 threads.  We measured Memcached's performance to be the
total time to complete 10,000 requests in each thread.

\subsection{Performance Evaluation}
\label{sec:eval:perf}

The main performance result of our work is that our designs impose performance
overheads that are low enough for use in deployed systems for many of our
experimental workloads.  Figure~\ref{fig:perfall} shows the slowdown suffered
by our benchmarks due to last writer slicing and \ctraps running with ``no-op''
trap hooks ({\em i.e.}, no handlers), relative to the natively executing
baseline.  In these experiments, we consider a run time overhead of 10\% or
less to be ideal for production use and a run time overhead of 50\% as
a reasonable upper bound on acceptable overhead for production use.

\paragraph{Last Writer Slices}
The data show that last writer slicing has very low overheads with a geometric
mean of less than 10\% across our server programs and less than 50\% across
PARSEC.  Such low overheads are likely to be acceptable in production.  In many
cases (Apache, MySQL, {\tt dedup}, {\tt canneal}), overheads are negligible.
In all but two cases ({\tt vips} \& {\tt swaptions}), the overhead of
collecting last writer slices is less than 100\%.

\paragraph{\ctraps}
\ctraps imposes a geometric mean overhead of 14\% for server applications and
110\% for PARSEC applications.  In 7 out of 13 of our tests, the overhead of
\ctraps is less than 50\%.  These seven low overhead benchmarks include all of
our server programs, as well as {\tt blackscholes}, {\tt dedup} and {\tt
canneal} from PARSEC.  The overhead for these applications is likely to be
tolerable in production.  {\tt vips} and {\tt swaptions} saw the highest
overheads -- around 400\%.  We discuss sources of high overheads in
Sections~\ref{sec:eval:conservative} and~\ref{sec:eval:parsecserver}.

By comparing \ctraps and last writer slicing, we see four applications ({\tt
dedup}, {\tt streamcluster}, {\tt fluidanimate}, {\tt ferret}) that have
\ctraps overhead that is probably too high for production use (averaging
153\%).  By comparison, those applications have a last writer slicing overhead
that is likely acceptable for production use (averaging 38\%).  The difference
indicates that these programs perform relatively more reads than other
applications, experiencing more overhead due to \ctraps's read instrumentation.

These high-level results support our main claim that the runtime overhead of
\ctraps is low enough for deployed systems for many applications, especially
for servers.  We now further analyze our observed overheads, illuminating
differences between the lower-overhead server programs and the higher-overhead
PARSEC programs.

\begin{figure}
\centering
\includegraphics[width=.9\columnwidth]{plots/perf.pdf}
\caption{\label{fig:perfall}Runtime overhead of \ctraps and Last Writer Slices (LWS).}
\end{figure}

\subsubsection{Overheads Due to Conservative Escape Analysis}
\label{sec:eval:conservative}
{\tt swaptions} had the highest run time overhead in our tests.  We
investigated the cause and determined it is largely due to shortcomings in our
escape analysis.  We located the inner-loop function {\tt
HJM\_SimPath\_Forward\_Blocking}.  The function uses two local matrices that
are allocated through an external call.  Our escape analysis cannot determine
that the matrices are only used locally in that function, so it conservatively
assumes the matrices escape.  We manually inlined the matrix allocation calls
(and a matching deallocation calls), and escape analysis eliminated their
instrumentation.  The program's overhead dropped from about 420\% overhead for
\ctraps to about 300\% overhead; the change for last writer slicing was from
300\% to about 130\% overhead.  We also looked at an output matrix passed into
the inner-loop function.  The matrix is allocated and deallocated in the
inner-loop's caller, but is never shared.  Eliminating instrumentation on
accesses to this matrix yielded an overhead of about 46\% for \ctraps, and to
just 30\% for last writer slicing.

These manual improvements to the escape analysis reduced the run time overheads
for these workloads from being unacceptable for any production environment to
being reasonable for use in production systems.  We expect a more sophisticated
escape analysis would find these local accesses and achieve these performance
gains automatically.

\subsubsection{PARSEC vs. Servers}
\label{sec:eval:parsecserver}
In our tests, server programs were better able to tolerate the overhead of
last writer slices and \ctraps.   We discern three reasons for this difference.

\paragraph{Independent Parallelism}
In most cases, the servers we evaluated have abundant, completely independent
parallel work.  Server programs use a pool of threads to handle requests.
Each thread incurs some instrumentation latency, but that latency is hidden by
other threads making progress on independent requests.  This analysis also
applies to the PARSEC applications that had performance similar to the servers.
For instance, {\tt blackscholes} uses largely independent threads to compute on
different regions of a matrix.  In addition, threads performing independent
computation ({\em i.e.}, non-communicating computation) are likely to operate
mostly on non-escaping, local data.  Local data are not instrumented by our
compiler, sparing the overhead.  We characterize the amount of local data in
each application in detail in Section~\ref{sec:char}.  We show that the server
programs have a larger proportion of provably local accesses than PARSEC.  The
difference supports the fact that the servers have more independent work ({\em
i.e.}, more local accesses) than PARSEC.

\paragraph{Latency Hiding I/O}
The servers perform more I/O.  Similarly to the effect of abundant parallelism,
I/O helps hide instrumentation latency.  While I/O is naturally part of many
applications, we aimed to minimize the effect of I/O in our experiments.  For
MySQL and Apache, we used local disks for logs and resources, and connected via
a local socket, rather than using the network.  In spite of these controls, I/O
remains, and hides added instrumentation latency in our server applications.
LevelDB's benchmark is designed to avoid all I/O, creating and manipulating an
in-memory database in a single process.  The absence of I/O may account for some
of the difference in overhead between LevelDB and the other servers.  Note that
the effects of I/O are not merely experimental noise -- in real deployed
software the latency hiding effects of I/O are likely to be even {\em more}
pronounced, favoring the use of our system.

\paragraph{Interference with Optimization}
The PARSEC programs are heavily hand-optimized. In several cases
resorting to hand-coded assembly and cache-aware memory access patterns.
Running instrumentation that accesses the \lwt alongside these computations may
disrupt their cache performance, or other delicately optimized behavior,
leading to higher overheads.

Such carefully optimized code is likely to have been written by expert
programmers.  As with other code written by experts ({\em e.g.}, lock-free
data-structures), this code may not require as much {\em in situ} analysis
support as code written by distributed teams of open source developers or
novice programmers.  Depending on the maturity and level of verification of a
piece of carefully hand-optimized code, it might make sense to elide 
last writer slicing and \ctraps code to preserve performance.

\subsection{Context-Sensitivity}
%{\bf TODO: Was this LWS or CTRAPS-NOOP or what?}
We briefly evaluated the cost of context-sensitivity using \ctraps with no-op
trap handlers.  With a bounded 2-address context, the average overhead for
server programs was 18\% and the average overhead for PARSEC was 145\%.  These
data show that a restricted form of context-sensitive analysis is possible in
\ctraps with overheads low enough for production.

\subsection{Evaluating CTraps Applications}
\label{sec:appperf}
We evaluated our CCI-Prev and communication graph collection CTraps
applications.  Our results show that these CTraps applications have modest
overheads that scale with their analysis complexity.  In several cases -- most
notably, in tests with our server benchmarks -- their overhead is low enough
for production use.

\subsubsection{Performance with \ctraps}

 Figure~\ref{fig:perfapps} shows the performance overheads of
our implementation as the slowdown with respect to the natively executing
baseline.  The overheads vary substantially across applications and the average
run time overheads are higher than for our baseline implementation
(Section~\ref{sec:eval:perf}).

In six cases (the four servers, {\tt blackscholes}, and {\tt dedup}), the overheads
for CCI-Prev implementations are tolerable for deployment use (around 50\%).
This result supports our claim that \ctraps helps build useful analyses that
have overheads low enough for deployment for an important class of
applications.

Overheads for communication graph collection are generally higher and probably
too high for production use in most instances.  The higher overheads are not a
negative result; instead, they illustrate the ``complexity proportional''
nature of \ctraps' overheads.  Our communication graph collection
implementation does more costly data structure manipulations on each trap, so
its overheads are higher.  \ctraps provides a low-overhead foundation and
applications scale their overhead as necessary.

%Of these cases, eliminating redundant read instrumentation
%(\ctrapsmm) improved performance in 5 cases.  For {\tt dedup} the eliminating
%redundant reads halved the performance overhead.  This result illustrates that
%for these applications, redundancy elimination provides additional performance
%benefit.

The PARSEC benchmarks have higher overhead than the servers for CCI-Prev and
graph collection.  Graph collection overheads for {\tt vips} and {\tt
streamcluster} are the worst cases we saw (23x and 47x, respectively).  We
described some reasons for the higher overheads in these tests in
Section~\ref{sec:eval:parsecserver}.  The overheads discussed there are further
exacerbated by the increased cost of reads in these applications.  These
overheads are too high for use in deployment;  however, we note that PARSEC
applications are batch focused and compute-intensive; we expect the
long-running, event-driven nature of server applications makes them a more
important target than PARSEC for these types of analyses in production.

{\bf TODO: Compare with these analyses implemented using binary instrumentation}

%In two of these cases, {\tt
%streamcluster} and {\tt swaptions}, we saw a substantial improvement in
%performance due to eliminating redundant read instrumentation ($~$4x for {\tt
%swaptions}).

%?The overheads we report are considerably lower than the overheads reported for
%?the compute-intensive benchmarks shown in the original CCI paper~\cite{cci}({\em
%?e.g.}LU Factorization \addtodo{BE SURE IT IS LU}).  While not efficient enough
%?to be deployed, these overheads are adequate for use in debugging.  Combining
%?\ctraps with sampling, as in CCI, is a promising route to deployable
%?efficiency.

\begin{figure}
\centering
\includegraphics[width=.9\columnwidth]{plots/appperf.pdf}
\caption{\label{fig:perfapps}Performance overheads of CCI-Prev and Graph Collection implemented with \ctraps.}
\end{figure}

\subsection{Characterizing \ctraps}
\label{sec:char}
We characterized the instrumentation pruned by escape analysis.  We report our
results in Table~\ref{tab:char}.  The table shows the percent of
instrumentation sites pruned and the absolute number of sites in parenthesis.

\begin{table}
\centering
\small
\begin{tabular}{l | r r }
\multirow{3}{*}{\bf Application} & \multicolumn{2}{c}{\bf \em Pruned } \\
               & \multicolumn{2}{c}{\bf \em Static Local}  \\ 
               & \multicolumn{2}{c}{\bf \em Operations}    \\ \hline

Apache         &  21.4\%&(12K)                    \\
MySQL          &  21.3\%&(53K)                    \\
memcached      &  65.8\%&(7.5K)                   \\
LevelDB        &  44.0\%&(4.5K)                   \\ \hline
blackscholes   &   7.5\%&(4)                      \\
dedup          &   22.0\%&(171)                   \\
canneal        &   28.6\%&(226)                   \\
x264           &   29.8\%&(5960)                  \\
vips           &   24.6\%&(9116)                  \\
ferret         &   20.9\%&(792)                   \\
fluidanimate   &   19.6\%&(149)                   \\
streamcluster  &   11.6\%&(47)                    \\
swaptions      &   8.3\% &(19)                    \\
\end{tabular}
\caption{\label{tab:char}Instrumentation pruned by static analysis.}
\end{table}

The data show that escape analysis can prove a substantial fraction of
operations are local and do not require instrumentation.  Across PARSEC,
between 8\% and 30\% of accesses were proved local.  In the server programs, a
larger proportion of accesses were able to be proved local (21\% -- 66\%).  The
difference is likely due to the fact that server threads perform more
independent work (as discussed in Section~\ref{sec:eval:parsecserver}).
Eliminating instrumentation using escape analysis is key to high performance.

{\bf TODO: Run without escape analysis?}

\section{Related Work}

There are several categories of prior work related to last writer slices and
\ctraps.

\paragraph{Program Slicing}
Slicing techniques~\cite{tipslicingsurvey} are techniques for identifying
useful subsequences of a program's instructions.   Slicing can work by
analyzing program code statically or analyzing execution traces dynamically.
Many slicing techniques discussed in ~\cite{tipslicingsurvey} monitor control
and data dependences to identify which parts of a program might be relevant to
a task ({\em e.g., debugging}).

Thin Slicing~\cite{thinslicing} is especially related to this work. Thin
Slicing monitors and reports a select subset of dependent operations from an
execution.  The goal is to eliminate some operations that would be included in
a traditional slice to make the slice simpler and less costly to collect.
While this work and Thin Slicing differ in what they collect and how, our
motivation to monitor a select subset of dependences was similar to theirs.

%is unlike our work in that our work focused on
%monitoring inter-thread dependences in production code without sampling or
%hardware support.

\paragraph{Debugging with Data Provenance Information}

Prior work has looked at using data provenance information in
debugging~\cite{whylineicse, whylinechi}.  The WhyLine debugger allowed
programmers to ask questions about why particular variables held particular
values during debugging.  WhyLine answered those questions by presenting data-
and control-flows that influenced those variables.  WhyLine showed that
provenance information is intuitive to programmers and makes debugging easier,
justifying our focus on provenance information in our last writer slicing
design.

There are several characteristics of this work that distinguish it from prior
work on debugging with data provenance information.  Unlike prior work, our
work focuses on isolating a limited subset of dependence information that we
can collect with run time overheads low enough for production use.
Furthermore, we focus on {\em concurrent} program debugging.  Prior work on
provenance debugging that focused mainly on single threaded programs and had
dependence slicing overheads are probably too high for production.

Bad Apple Tracking~\cite{badapples} is also a form of debugging with data
provenance information. This work records the {\em origin} of an unusable value
and can report it to a programmer when it is used in a way that causes a
failure.

\paragraph{Communication Tracking in Concurrent Programs}
Several techniques have been proposed do some form of dependence and
communication tracking to analyze concurrent programs.

CCI~\cite{cci}, Bugaboo~\cite{bugaboo}, Recon~\cite{recon}, and
DefUse~\cite{defuse} all explicitly track inter-thread memory dependences at
runtime.  Unlike our work, several of these techniques were not designed to run
in production systems~\cite{recon,defuse}.  A distinguishing characteristic of
our work is that it is effective and fast enough for production without the
need to aggregate information from multiple different sampled executions, like
CCI~\cite{cci}.  Additionally, our system does not require invasive hardware
changes like Bugaboo~\cite{bugaboo} or LBA~\cite{paralog}.

Other techniques~\cite{threadclustering,threadcriticality} use  hardware debug
registers to monitor dependences efficiently enough for production use.  These
techniques are similar to our work in that they track how threads interact.
These techniques are different from our work in that they track information
specific to the performance and our system provides general support.

\paragraph{Extensible Program Instrumentation}
There are many systems for building analyses using extensible program
instrumentation.  Binary instrumentation
systems~\cite{pin,dynamorio,valgrind,roadrunner} provide completely general
support to instrument arbitrary code -- instrumenting a program to track
communciation is possible in such systems.

These systems are similar to our work in that they aim to enable dynamic
analyses and runtime systems.  They are different from our work because they
rely on dynamic rewriting, yielding overheads often too high for deployment.
Instead, we trade generality for performance. Last writer slices and \ctraps
provide extensible support in the restricted domain of tracking communication
with overheads low enough for production.  Another difference is that our work
handles some of the delicate, performance sensitive code required to track
communication.  In general instrumentation systems, programmers write that code
from scratch at their own risk.

\section{Conclusions}
This work observes that we can collect a useful form of data provenance
information with overheads low enough for use in deployed systems.

We developed system support that collects last writer slices with overheads
well within an acceptable range for production.  We showed that, during
debugging, the provenance information provided by last writer slices reveals
crucial connections between different threads' otherwise disparate regions of
code.  Understanding how code in different threads interacts is essential to
concurrency debugging and this work helps with that understanding.

Using last writer slices, we then built \ctraps, an extensible framework for
implementing concurrent program analyses that interpose on inter-thread
communication.  Trap handlers implement such analyses in just a few lines of
code.  We showed that it is possible to build useful analyses using \ctraps
that are efficient enough for deployment use in an important set of programs
({\em e.g.}, servers, key-value stores, databases, {\em etc.}).

An interesting direction for future work is to evaluate the role of minimal
hardware support for \ctraps.  We will also further study the impact of
provenance information on debugging in both sequential and concurrent
applications.

\bibliography{bib}{}
\bibliographystyle{abbrv}

\end{document}